### 2021-04-01 homology.tex

\subsection{Homology theories}

\begin{itemize}
	\item Definition
	\item Examples, main: singular with integer coeff and cech with field 
	\item Mayer-Vietoris
\end{itemize}