### 2021-04-01 preliminaries.tex

\section{Preliminaries}

\input{filtrations}
\input{homology}
\input{persistence}