### 2021-04-03 homology.tex
Consolidate sections connectivity, counterexample and theorem

\section{Homology theories}

\begin{itemize}
	\item Definition
	\item Examples, main: singular with integer coeff and cech with field 
	\item Mayer-Vietoris
\end{itemize}