### 2021-04-09 homology.tex

\section{Homology theories}

\begin{itemize}
\item Definition (generalized homology theory, without dimension axiom, but with Milnor's additivity axiom)
\item Examples, main: singular with integer coeff and cech with field \cite{Kelly.1961}
\item mention Dowker duality
\item Mayer-Vietoris
\end{itemize}

\todo{Uli: I saw a discussion about the topological invariance of Vietoris homology for non-compact metric spaces.
I thought that it suggested that invariance holds only for compact spaces,
but I can't find it anymore.

What I found was this:
\url{http://pldml.icm.edu.pl/pldml/element/bwmeta1.element.zamlynska-8f369e96-b1eb-4ec9-99e2-8c366367a5a4} claims invariance also for non-compact.
\url{https://www.ams.org/journals/tran/1947-062-02/S0002-9947-1947-0024128-6/S0002-9947-1947-0024128-6.pdf} mentions a non-invariant property "null-bounding"
}