### 2021-04-09 homology.tex
Add section by section to intro

\section{Homology theories} \label{s:homology}

\begin{itemize}
	\item Definition
	\item Examples, main: singular with integer coeff and cech with field 
	\item Mayer-Vietoris
\end{itemize}