### 2021-06-15 main.tex
Change doc-class to springer and modify abstract

%\immediate\write18{bibtex \jobname}
\documentclass[nospthms, a4paper, final]{svjour3}
\setlength{\textwidth}{\dimexpr\pdfpagewidth-2in}
\setlength{\textheight}{\dimexpr\pdfpageheight-2in}
\usepackage{mathptmx}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{tikz-cd}
\usepackage[numbers,sort,compress]{natbib}
\usepackage{graphicx}
\usepackage{csquotes}
\usepackage[bookmarks=true,
bookmarksnumbered=true, breaklinks=true,
pdfstartview=FitH, hyperfigures=true,
plainpages=false, naturalnames=true,
colorlinks=true, pagebackref=true,
pdfpagelabels]{hyperref}
\hypersetup{
	linktocpage,
	colorlinks,
	citecolor=blue,
	linkcolor=blue,
	urlcolor=blue}
\usepackage[capitalize, noabbrev]{cleveref}
\usepackage{todo}

% theorems
\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\Crefname{thm}{Theorem}{Theorems}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\Crefname{prop}{Proposition}{Propositions}
\newtheorem{claim}[thm]{Claim}
\newtheorem{conj}[thm]{Conjecture}
\Crefname{conj}{Conjecture}{Conjectures}
\theoremstyle{definition}
\newtheorem{defi}[thm]{Definition}
\newtheorem{ex}[thm]{Example}
\newtheorem{rem}[thm]{Remark}

% commands
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\HS}{\mathrm{HS}}
\newcommand{\HT}{\mathrm{HT}}
\newcommand{\HLC}{\mathrm{HLC}}
\newcommand{\piT}{\pi\mathrm{T}}
\newcommand{\LC}{\mathrm{LC}}
\newcommand{\NH}{\mathrm{NH}}
\newcommand{\piLC}{\pi\mathrm{LC}}
\newcommand{\HE}{\mathbb{H}^d}
\renewcommand{\H}{\widetilde{H}}
\DeclareMathOperator{\interior}{int}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\rad}{rad}
\newcommand\CH{\check{H}}
\newcommand{\Mod}[1]{#1-\mathbf{Mod}}
\newcommand{\PFD}{\mathrm{PFD}}
\newcommand{\PFG}{\mathrm{PFG}}
\DeclareMathOperator{\Nrv}{Nrv}
\DeclareMathOperator{\Cov}{Cov}
\DeclareMathOperator{\Vietoris}{Vtr}
\DeclareMathOperator{\Balls}{Balls}
\DeclareMathOperator*{\colim}{colim}
\DeclareMathOperator{\crit}{crit}
\newcommand\MCH{M\check{H}}
\newcommand\VH{VH}
\newcommand\MVH{MVH}
\newcommand{\obs}{\dashrightarrow}
\newcommand{\multiplicityDomain}{\mathcal{E}}
\newcommand{\Top}{\mathsf{Top}}
\newcommand{\Vect}{\mathsf{Vect}}
\newcommand{\anibal}[1]{\textcolor{blue}{\underline{Anibal}: #1}}

\begin{document}

\title{Persistence in functional analysis and a correction to a theorem of Morse}

\author{author1 \and author2 \and author3}

\institute{author1 \at where \\ \email{@} \and
	author2 \at where \\ \email{@} \and
	author3 \at where \\ \email{@} \and
}

\authorrunning{author1, author2, author3}

\maketitle

\begin{abstract}
	During the 1930's Morse developed the theory of functional topology relating the critical set of a semi-continuous functional and the topology of its sublevel sets.
	Morse and Tompkins applied this body of work to prove the Unstable Minimal Surface Theorem in the setting defined by Douglas' solution to Plateau's Problem.
	In recent years, persistence homology has led to fast passed developments in both mathematical theory and practice within several fields including simplectic geometry, WHAT ELSE? and neuroscience. In this article, we present a modern account of Morse's functional topology from the point of view of persistence theory.
	We do so to allow for future applications of persistence techniques in functional analysis, as well as to correct a gap we identified in the proof of the Unstable Minimal Surface Theorem of Morse and Tompkins.
\end{abstract}

\tableofcontents
\input{introduction}
\input{persistence}
\input{inequalities}
\input{connectivity}
\input{surfaces}
\appendix
\input{vietoris}

\section*{Acknowledgements}
This research has been supported by German Research Foundation (DFG) through the Collaborative Research Center SFB/TRR 109 \emph{Discretization in Geometry and Dynamics}, the Collaborative Research Center SFB/TRR 191 \emph{Symplectic Structures in Geometry, Algebra and Dynamics}, the Cluster of Excellence EXC-2181/1 \emph{STRUCTURES}, and the Research Training Group RTG 2229 \emph{Asymptotic Invariants and Limits of Groups and Spaces}.

A.M-M. acknowledges financial support from Innosuisse grant \mbox{32875.1 IP-ICT - 1} and the hospitality of the Laboratory for Topology and Neuroscience at EPFL where part of this work developed.

\bibliographystyle{abbrvnaturl}
\bibliography{biblio}

\end{document}