### 2021-07-22 main.tex

\documentclass[nospthms, a4paper, final]{svjour3}
\setlength{\textwidth}{\dimexpr\pdfpagewidth-2in}
\setlength{\textheight}{\dimexpr\pdfpageheight-2.4in}
\usepackage{mathptmx}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{tikz-cd}
\usepackage[numbers,sort,compress]{natbib}
\usepackage{graphicx}
\usepackage{csquotes}
\usepackage[bookmarks=true,
bookmarksnumbered=true, breaklinks=true,
pdfstartview=FitH, hyperfigures=true,
plainpages=false, naturalnames=true,
colorlinks=true, pagebackref=true,
pdfpagelabels]{hyperref}
\hypersetup{
	linktocpage,
	colorlinks,
	citecolor=blue,
	linkcolor=blue,
	urlcolor=blue}
\usepackage[capitalize, noabbrev]{cleveref}
\usepackage{todo}

% theorems
\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\Crefname{thm}{Theorem}{Theorems}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\Crefname{prop}{Proposition}{Propositions}
\newtheorem{claim}[thm]{Claim}
\newtheorem{conj}[thm]{Conjecture}
\Crefname{conj}{Conjecture}{Conjectures}
\theoremstyle{definition}
\newtheorem{defi}[thm]{Definition}
\newtheorem{ex}[thm]{Example}
\newtheorem{rem}[thm]{Remark}

% commands
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\HS}{\mathrm{HS}}
\newcommand{\HT}{\mathrm{HT}}
\newcommand{\HLC}{\mathrm{LHS}}
\newcommand{\piT}{\pi\mathrm{T}}
\newcommand{\LC}{\mathrm{LC}}
\newcommand{\NH}{\mathrm{NH}}
\newcommand{\piLC}{\pi\mathrm{LC}}
\newcommand{\HE}{\mathbb{H}^d}
\renewcommand{\H}{H}
\DeclareMathOperator{\interior}{int}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\rad}{rad}
\newcommand\CH{\check{H}}
\newcommand{\Mod}[1]{#1-\mathbf{Mod}}
\newcommand{\PFD}{\mathrm{PFD}}
\newcommand{\PFG}{\mathrm{PFG}}
\DeclareMathOperator{\Nrv}{Nrv}
\DeclareMathOperator{\Cov}{Cov}
\DeclareMathOperator{\Vietoris}{Vtr}
\DeclareMathOperator{\Balls}{Balls}
\DeclareMathOperator*{\colim}{colim}
\DeclareMathOperator{\crit}{crit}
\newcommand\MCH{M\check{H}}
\newcommand\VH{VH}
\newcommand\MVH{MVH}
\newcommand{\obs}{\dashrightarrow}
\newcommand{\multiplicityDomain}{\mathcal{E}}
\newcommand{\Top}{\mathsf{Top}}
\newcommand{\Vect}{\mathsf{Vect}}
\renewcommand{\th}{^{\mathrm{th}}}
\newcommand{\anibal}[1]{\textcolor{blue}{\underline{Anibal}: #1}}

\begin{document}

\title{Persistence in functional topology and a correction to a theorem of Morse}

\author{Ulrich Bauer \and Anibal M.~Medina-Mardones \and Maximilian~Schmahl}

\institute{
    Ulrich Bauer \at Technical University of Munich \\ \email{mail@ulrich-bauer.org} \and
	Anibal M.~Medina-Mardones \at University of Notre Dame du Lac \at Max Planck Institute for Mathematics in Bonn \\ \email{ammedmar@mpim-bonn.mpg.de} 
	\and
	Maximilian~Schmahl \at Heidelberg University \\ \email{mschmahl@mathi.uni-heidelberg.de}
}

\authorrunning{U.~Bauer, A.~Medina-Mardones, M.~Schmahl}

\maketitle

\begin{abstract}
	During the 1930s, Marston Morse developed a vast generalization of what is commonly known as Morse theory relating the critical points of a semi-continuous functional with the topology of its sublevel sets.
	Morse and Tompkins applied this body of work, referred to as functional topology, to prove the Unstable Minimal Surface Theorem in the setting defined by Douglas' solution to Plateau's Problem.
	Several concepts introduced by Morse in this context can be seen as early precursors to the theory of persistent homology, which by now has established itself as a popular tool in applied and theoretical mathematics.
	In this article, we provide a modern redevelopment of the homological aspects of Morse's functional topology from the perspective of persistence theory.
	We adjust several key definitions and prove stronger statements, including a generalized version of Morse inequalities, in order to allow for novel uses of persistence techniques in functional analysis and symplectic geometry.
	As an application, we identify and correct a mistake in the proof of the Unstable Minimal Surface Theorem of Morse and Tompkins.
\end{abstract}

\tableofcontents
\input{introduction}
\input{persistence}
\input{inequalities}
\input{connectivity}
\input{surfaces}
\appendix
\input{vietoris}

\section*{Acknowledgements}
This research has been supported by German Research Foundation (DFG) through the Collaborative Research Center SFB/TRR 109 \emph{Discretization in Geometry and Dynamics}, the Collaborative Research Center SFB/TRR 191 \emph{Symplectic Structures in Geometry, Algebra and Dynamics}, the Cluster of Excellence EXC-2181/1 \emph{STRUCTURES}, and the Research Training Group RTG 2229 \emph{Asymptotic Invariants and Limits of Groups and Spaces}.

A.M-M. acknowledges financial support from Innosuisse grant \mbox{32875.1 IP-ICT-1} and the hospitality of the Laboratory for Topology and Neuroscience at EPFL, where part of this work developed.

\bibliographystyle{abbrvnaturl}
\bibliography{biblio}

\end{document}