### 2021-09-01 func_top.tex

\documentclass{amsart}
\input{aux/style}
\input{aux/commands}

\title[Persistence in functional topology]{Persistence in functional topology and a correction to a theorem of Morse}

\author[U. Bauer]{Ulrich Bauer}
\address{U.~Bauer, Technical University of Munich}
\email{\href{mailto:mail@ulrich-bauer.org}{mail@ulrich-bauer.org}}

\author[A. Medina-Mardones]{Anibal M. Medina-Mardones}
\address{A.~Medina-Mardones, Max Planck Institute for Mathematics \and University of Notre Dame}
\email{\href{mailto:ammedmar@mpim-bonn.mpg.de}{ammedmar@mpim-bonn.mpg.de}}

\author[M. Schmahl]{Maximilian Schmahl}
\address{M.~Schmahl, Heidelberg University}
\email{\href{mailto:mschmahl@mathi.uni-heidelberg.de}{mschmahl@mathi.uni-heidelberg.de}}

\subjclass[2020]{Primary: 55N31, 58E05. Secondary: 58E12, 54D05, 55N05.}

\date{\today}

\begin{document}
	\input{aux/abstract}
	\maketitle
	\input{sec/introduction}
	\input{sec/persistence}
	\input{sec/inequalities}
	\input{sec/connectivity}
	\input{sec/surfaces}
	\appendix
	\input{sec/vietoris}
	\input{aux/acknowledgment}
	\bibliographystyle{aux/abbrvnaturl}
	\bibliography{aux/biblio}
\end{document}