### 2021-10-12 aux/abstract.tex
After interchanging summary and abstract

\begin{abstract}
The primary contribution of this work consists in the introduction of topological conditions on a broad class of filtered spaces ensuring their associated persistent homology modules admit persistence diagrams, which, in particular, imply that they satisfy generalized Morse inequalities.
We illustrate the applicability of this result by recasting the original proof of the Unstable Minimal Surface Theorem given by Morse and Tompkins in a modern and rigorous framework.
\end{abstract}