### 2021-10-12 func_top.tex

\immediate\write18{biber \jobname}

\documentclass{amsart}
\include{aux/style}
\include{aux/commands}

\title[Persistence in functional topology]{Persistence in functional topology:
existence of persistence diagrams 
and
Morse inequalities
}

%from local connectivity
%to 
%q-tameness
%to 
%persistence diagram
%to
%Morse inequalities

%q-tameness,
%Morse inequalities,
%and
%existence of persistence diagrams

%Morse inequalities,
%q-tameness,
%and
%existence of persistence diagrams

%generalized Morse inequalities,
%and
%existence of persistence diagrams

%Local connectivity, q-tameness, and persistence diagrams

\author[U.~Bauer]{Ulrich Bauer}
\address{U.B., Technical University of Munich}
\email{\href{mailto:mail@ulrich-bauer.org}{mail@ulrich-bauer.org}}

\author[A.~Medina-Mardones]{Anibal M. Medina-Mardones}
\address{A.M-M., Max Planck Institute for Mathematics \and University of Notre Dame}
\email{\href{mailto:ammedmar@mpim-bonn.mpg.de}{ammedmar@mpim-bonn.mpg.de}}

\author[M.~Schmahl]{Maximilian Schmahl}
\address{M.S., Heidelberg University}
\email{\href{mailto:mschmahl@mathi.uni-heidelberg.de}{mschmahl@mathi.uni-heidelberg.de}}

\subjclass[2020]{55N31, 58E05, 58E12, 54D05, 55N05.}

\date{\today}

\begin{document}
	\input{aux/abstract}
	\maketitle
	\input{sec/introduction}
	\input{sec/persistence}
	\input{sec/inequalities}
	\input{sec/connectivity}
	\input{sec/surfaces}
	\appendix
	\input{sec/vietoris}
	\input{aux/acknowledgment}
	\printbibliography
\end{document}