### 2021-10-14 sec/abstract.tex
After spliting into ss Sect 5

%!TEX root = ../func_top.tex

\begin{abstract}
We introduce topological conditions on a broad class of filtered spaces ensuring that their associated persistent homology modules admit persistence diagrams, which, in particular, implies that they satisfy generalized Morse inequalities.
We illustrate the applicability of these results by recasting the original proof of the Unstable Minimal Surface Theorem given by Morse and Tompkins in a modern and rigorous framework.
\end{abstract}