### 2022-03-28 referee_reply.tex
Add referee_reply file and first fixes

\documentclass{article}
\usepackage{csquotes}
\title{Author's reply to the referee report of \\
\textsc{``PERSISTENCE IN FUNCTIONAL TOPOLOGY: EXISTENCE OF PERSISTENCE DIAGRAMS AND MORSE INEQUALITIES''}
}
\begin{document}
	\maketitle
	\section{First point.} TBW

	\section{Second point.} TBW

	\section{Third point.} TBW

	\section{Additional suggestion.} TBW

	\section{List of typos, minor remarks and suggestions.}
	\begin{enumerate}
		\item Replaced as suggested (RAS), thank you.
		\item RAS, thank you.
		\item We did mean to ``functional analysis''. Thank you for the question.
		\item RAS, thank you.
		\item RAS, thank you.
		\item Before discussing more details of the stability result, including the concepts related to the metric theory of persistence modules we added the sentence:
		\begin{quote}
			We now provide more details regarding this result although they are not used in the present work.
		\end{quote}
		\item Definition moved as suggested.
		\item RAS, thank you.
		\item RAS, thank you.
		\item RAS, thank you.
		\item RAS, thank you. (ULI, PLEASE CHECK THIS CHANGE)
		\item RAS, thank you.
		\item (MAX, PLEASE ADDRESS THIS ONE)
		\item RAS, thank you.
		\item Replaced the preceding paragraph with:
		\begin{quote}
			In analogy to Morse's definitions, we may define, for an integer $d$ and real numbers $t$ and $\epsilon > 0$, the $(d, t, \epsilon)$-\emph{cap number} of our graded q-tame persistence module $M$ in terms of its persistence diagram as:
		\end{quote}
		(MAX, PLEASE CHECK THIS CHANGE)
		\item MAX, I REMEMBER ASKING YOU ABOUT THIS. WHAT WAS YOUR ANSWER? COULD YOU TAKE CARE OF THIS POINT?
		\item RAS, thank you.
		\item RAS, thank you.
		\item RAS, thank you.
		\item TBW
		\item RAS, thank you.
		\item (I THINK WE ARE ASSUMING HAUSDORFF, RIGHT? IS THIS NOT MENTIONED?)
		\item TBW
		\item TBW
		\item RAS, thank you.
		\item (I DO NOT KNOW)
		\item RAS, thank you.
		\item RAS, thank you.
		\item (I DO NOT KNOW)
		\item TO BE ADDRESS AFTER DISCUSSING WHAT ``homotopically critical point'' ARE.
		\item (I DO NOT KNOW)
		\item RAS, thank you.
		\item RAS, thank you.
		\item RAS, thank you.
		\item RAS, thank you.
		\item (I DO NOT KNOW)
		\item TBW
		\item (I DO NOT KNOW)
		\item RAS, thank you.
		\item TBW
	\end{enumerate}
\end{document}