### 2022-10-12 ref_reply3.tex
Add ref reply 3

\documentclass{article}
\usepackage{csquotes}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}

\newcommand{\cc}{\check{c}}

\title{Author's third reply to the referee report of \\ \textsc{
		``PERSISTENT HOMOLOGY FOR FUNCTIONALS''
	}
}
\author{Bauer \and Medina-Mardones \and Schmahl}
\begin{document}
	\maketitle
	We want to thank the referee for his careful reading of our revised article, again providing very helpful commentary and uncovering many of our inaccuracies.

	\section{First point.}
	TBW
	\section{Second point.}
	TBW
	\section{third point.}
	TBW
	\section{List of typos, minor remarks and suggestions.}
	\begin{enumerate}
		\item TBW
		\item Changed as suggested (A.S.).
		\item Changed as suggested (A.S.).
		\item TBW
		\item TBW
		\item Changed as suggested (A.S.).
		\item TBW
		\item TBW
		\item TBW
		\item TBW
		\item TBW
		\item TBW
		\item TBW
	\end{enumerate}

	\section{Other changes}
	TBW
\end{document}