### 2023-05-16 sec/acknowledgment.tex

%!TEX root = ../func_top.tex

\section*{Acknowledgements}
The authors wish to thank an anonymous referee for his extensive feedback and valuable suggestions, which have greatly improved the present work.

U.B. has been supported by the German Research Foundation (DFG) through the Collaborative Research Center SFB/TRR 109 \emph{Discretization in Geometry and Dynamics} -- 195170736.
A.M-M. acknowledges financial support from Innosuisse grant \mbox{32875.1 IP-ICT-1} and the hospitality of the Laboratory for Topology and Neuroscience at EPFL.
M.S. has been supported by the German Research Foundation (DFG) through the Collaborative Research Center SFB/TRR 191 \emph{Symplectic Structures in Geometry, Algebra and Dynamics} -- 281071066, the Cluster of Excellence EXC 2181 \emph{STRUCTURES} -- 390900948, and the Research Training Group RTG 2229 \emph{Asymptotic Invariants and Limits of Groups and Spaces} -- 281869850.