### 2023-10-10 ref_reply4.tex
Fix typos found by referee (round 4)

\documentclass{article}
\usepackage{csquotes}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}

\newcommand{\cc}{\check{c}}

\title{Author's fourth reply to the referee report of \\ \textsc{
		``PERSISTENT HOMOLOGY FOR FUNCTIONALS''
	}
}
\author{Bauer \and Medina-Mardones \and Schmahl}
\begin{document}
	\maketitle

	\section*{Corrections}
	\begin{enumerate}
		\item Fixed.
		\item Fixed.
		\item ...
		\item Fixed.
		\item ...
		\item ...
		\item Fixed.
		\item ...
		\item ...
		\item Fixed
	\end{enumerate}
\end{document}